---
title: "Analysen aus Praesentation"
format:
  html:
    toc: true
    number-sections: true
    highlight-style: github
editor: visual
---

## Quarto

```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(haven, psych,
               sjmisc, sjPlot, writexl, lmerTest,
               tidyverse, multilevelTools, parameters)
```

## Sitzung 3

### MLM-Motivierendes Beispiel

```{r}
load("../data/df_example1.RData")
load("../data/df_example1c.RData")

df_example1d <- read.table(file = "../data_simulation/example1drep2.dat", header = FALSE, 
        na.strings = "*", strip.white = TRUE) |> as_tibble()
names(df_example1d) <- c("w", "y", "x", "id")
df_example1d <- df_example1d |> select(id, everything())

df_example1d <- df_example1d |>
  mutate(id = as.factor(id)) |> 
  de_mean(y, x, grp = "id")
```

### Variante 1: Klassisches lineares Modell ohne Ber체cksichtigung der hierarchischen Datenstruktur

```{r}
v1_results <- lm(y ~ x, df_example1d)
model_parameters(v1_results) |> print_html()

df_example1d |> 
    mutate(id = as.factor(id)) |> 
    ggplot(aes(x = x, y = y)) + 
    geom_point(aes(color = id)) +
    geom_smooth(method = lm)
ggsave("Plots/01_standardlm_idseingef채rbt.png", width = 10, height = 7)

df_example1d |> 
    mutate(id = as.factor(id)) |> 
    ggplot(aes(x = x, y = y)) + 
    geom_point() +
    geom_smooth(method = lm)
ggsave("Plots/01a_standardlm_idsgrau.png", width = 7, height = 7)

# zeigt nur 50 - der 체bersicht halber
```

### Variante 2: Personen-Mittelwerte

```{r}
df_personlevel <- df_example1d |> group_by(id) |> 
  summarise(across(c(y,x), mean))

v2_results <- lm(y ~ x, df_personlevel)
model_parameters(v2_results) |> print_html()
```

Grossere Standardfehler / Konfidenzintervalle

### Variante 3: Mehrebenen-Modell

```{r}
ri.fs_modell <- lmer(y ~ x_dm + x_gm + (1 | id), data = df_example1d)
model_parameters(ri.fs_modell) |> print_html()


df_example1d |>
  filter(id %in% 1:20) |> 
    mutate(id = as.factor(id)) |> 
    ggplot(aes(x = x, y = y)) + 
    geom_point(aes(color = id)) +
    geom_smooth(method = lm, se = FALSE, size = 1.5) +
    geom_smooth(aes(color = id), method = lm, se = FALSE, alpha = 0.2)

ggsave("Plots/01_mlm_individuelleregression.png", width = 10, height = 7)


df_example1d_withpred <- df_example1d %>%
  mutate(predicted = predict(ri.fs_modell))

# Plot raw data and regression line
df_example1d_withpred |> 
  filter(id %in% 1:4) |> 
  ggplot(aes(x = x, y = y, color = id)) +
  geom_point() +  # Raw data points
  geom_line(aes(y = predicted), size = 1.2) +  # Regression line from the model
  labs(title = "Multilevel Model with Raw Data and Regression Line",
       x = "X Variable",
       y = "Y Variable") +
    facet_wrap(vars(id))
ggsave("Plots/01_fixedsloperandomintercept.png", width = 10, height = 7)

```

```{r}
ri.rs_modell <- lmer(y ~ x_dm + x_gm + (x_dm + 1 | id), data = df_example1d)
df_example1d_withpredrs <- df_example1d %>%
  mutate(predicted = predict(ri.rs_modell))
# Plot raw data and regression line
df_example1d_withpredrs |> 
  filter(id %in% 1:12) |> 
  ggplot(aes(x = x, y = y, color = id)) +
  geom_point() +  # Raw data points
  geom_line(aes(y = predicted), size = 1.2) +  # Regression line from the model
  labs(title = "Multilevel Model with Raw Data and Regression Line",
       x = "X Variable",
       y = "Y Variable") +
    facet_wrap(vars(id))

ggsave("Plots/01_randomsloperandomintercept.png", width = 10, height = 7)
```

Separat pro Person

```{r}
df_example1d |>
    filter(id %in% 1:4) |>
  mutate(time = rep(1:10, 4)) |> 
    mutate(id = as.factor(id)) |> 
    ggplot(aes(x = time, y = y, group = id, color = id)) + 
    geom_point() +
    geom_smooth(method = lm) +
    facet_wrap(vars(id))
ggsave("Plots/01_separat_zeit.png", width = 10, height = 7)
```

```{r}
sep_plot <- function(df, y, x, idvar, idnum) {
  df |>
    filter(!!sym(idvar) %in% idnum) |> 
    mutate(!!sym(idvar) := as.factor(!!sym(idvar))) |> 
    ggplot(aes(x = x, y = y, group = !!sym(idvar), color = !!sym(idvar))) + 
    geom_point() +
    geom_smooth(method = lm) +
    facet_wrap(vars(!!sym(idvar)))
}

sep_plot(df_example1d, "y", "x", "id", 1:4)
ggsave("Plots/01_separat_yx.png", width = 10, height = 7)
```

## Sitzung 4

### Standard LM

```{r}
data(diamonds)
diamonds_z <- diamonds |> sjmisc::std()
v1_results <- lm(price_z ~ carat_z, diamonds_z)
v1_results
p <- diamonds_z |> 
    ggplot(aes(x = carat_z, y = price_z)) + 
    geom_smooth(method = lm, color = "black", se = FALSE) +
   theme_minimal() +
 scale_x_continuous(breaks = function(x) floor(min(x)):ceiling(max(x))) +  # Auto integer breaks
  scale_y_continuous(breaks = function(y) floor(min(y)):ceiling(max(y))) # Auto integer breaks

ggsave("plots/diamonds.png", plot = p, device = "png", width = 6, height = 4, dpi = 300)

```

### Separate Plots pro Person f체r Sitzung 4

```{r}
load("../data/df_example1.RData")
load("../data/df_example1c.RData")
nullmodel <- lmer(y ~ (1 | id), data = df_example1)
df_example1 <- df_example1 |> mutate(ypred_null = predict(nullmodel))

df_example1 |>
  filter(id %in% 1:4) |> mutate(time = rep(1:10, 4)) |>
  mutate(id = as.factor(id)) |>
  ggplot(aes(
    x = x,
    y = y,
    group = id,
    color = id
  )) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = ypred_null), size = 1) +
              facet_wrap(vars(id))

ggsave("Plots/separat_nullmodel.png", width = 10, height = 7)
```

```{r}
ri.fs_modell <- lmer(y ~ x_dm + (1 | id), data = df_example1)
df_example1 <- df_example1 |> mutate(ypred_rifs = predict(ri.fs_modell))

df_example1 |>
  filter(id %in% 1:4) |> mutate(time = rep(1:10, 4)) |>
  mutate(id = as.factor(id)) |>
  ggplot(aes(
    x = x,
    y = y,
    group = id,
    color = id
  )) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = ypred_rifs), size = 1) +
              facet_wrap(vars(id))

ggsave("Plots/separat_rifs.png", width = 10, height = 7)
```

```{r}
ri.rs_modell <- lmer(y ~ x_dm + (1 + x_dm | id), data = df_example1)

df_example1 <- df_example1 |> mutate(ypred_rirs = predict(ri.rs_modell))

df_example1 |>
  filter(id %in% 1:4) |> mutate(time = rep(1:10, 4)) |>
  mutate(id = as.factor(id)) |>
  ggplot(aes(
    x = x,
    y = y,
    group = id,
    color = id
  )) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = ypred_rirs), size = 1) +
              facet_wrap(vars(id))
  

ggsave("Plots/separat_rirs.png", width = 10, height = 7)

df_example1 |>
  filter(id %in% 1:4) |> mutate(time = rep(1:10, 4)) |>
  mutate(id = as.factor(id)) |>
  ggplot(aes(
    x = x,
    y = y,
    group = id,
    color = id
  )) +
  geom_point(alpha = 0.5) +
  geom_line(aes(y = ypred_rirs), size = 1) +
  geom_line(aes(y = ypred_rifs), size = 1, color = "black") +
              facet_wrap(vars(id))

ggsave("Plots/separat_rirs_withfs.png", width = 10, height = 7)

```

## Sitzung 5

Interaktionsmuster: Puffer

```{r}
set.seed(123)  # For reproducibility

# Generate predictors
n <- 500
X <- rnorm(n, mean = 0, sd = 1)  # Independent variable
W <- rnorm(n, mean = 0, sd = 1)  # Moderator

# Define coefficients
b_X <- -0.5    # Negative effect of X on Y at low W
b_W <- 0.3     # Effect of W on Y
b_XW <- 0.6    # Positive interaction effect (buffering)

# Generate outcome variable
Y <- 2 + (b_X * X) + (b_W * W) + (b_XW * X * W) + rnorm(n, mean = 0, sd = 1)

# Create dataframe
data <- data.frame(X, W, Y)

# Inspect the data
summary(lm(Y ~ X * W, data = data))

# Load necessary library
library(sjPlot)

# Fit the model
model <- lm(Y ~ X * W, data = data)

# Create an interaction plot
plot_model(model, type = "int", terms = c("X", "W"), mdrt.values = "meansd")
ggsave("Plots/session5_bufferinteraction.png")
```

```{r}
set.seed(123)  # For reproducibility

# Generate predictors
n <- 500
X <- rnorm(n, mean = 0, sd = 1)  # Independent variable
W <- rnorm(n, mean = 0, sd = 1)  # Moderator

# Define coefficients
b_X <- 0.3    # Negative effect of X on Y at low W
b_W <- 0.2     # Effect of W on Y
b_XW <- 0.5    # Positive interaction effect (buffering)

# Generate outcome variable
Y <- 2 + (b_X * X) + (b_W * W) + (b_XW * X * W) + rnorm(n, mean = 0, sd = 1)

# Create dataframe
data <- data.frame(X, W, Y)

# Inspect the data
summary(lm(Y ~ X * W, data = data))

# Load necessary library
library(sjPlot)

# Fit the model
model <- lm(Y ~ X * W, data = data)

# Create an interaction plot
plot_model(model, type = "int", terms = c("X", "W"), mdrt.values = "meansd")
ggsave("Plots/session5_synergyinteraction.png")
```

## Sample Paper

### Einlesen

```{r}
df_example_cli <- read.table(file = "../data_simulation/example_cli1.dat", header = FALSE, 
        na.strings = "999.000000", strip.white = TRUE) |> as_tibble()
names(df_example_cli) <- c("w", "y", "x", "pattern", "id")
df_example_cli <- df_example_cli |> select(id, y, x, w) |> 
  rename(illtask = x, negativ = y, support = w)

df_example_cli <- df_example_cli |> 
  de_mean(illtask, negativ, grp = "id")

df_example_cli <- df_example_cli |> 
  group_by(id) |> 
  mutate(
    negativ_dm_l1 = lag(negativ_dm),
    day = 1:10
  )

df_example_cli <- df_example_cli |> filter(!is.na(negativ)) |> ungroup()

df_example_cli <- df_example_cli |>
  sjlabelled::var_labels(
  id = "Personen-ID",
  illtask = "Daily Illegitimate Tasks",
  negativ = "Daily Negative Affect",
  negativ_gm = "Neg. Affect (person average)",
  negativ_dm = "Neg. Affect (person-mean centered)",
  negativ_dm_l1 = "Previous-day Neg. Affect",
  illtask_dm = "Illegitimate Tasks (person-mean centered)",
  illtask_gm = "Illegitimate Tasks (person average)",
  support = "Coworker support",
  day = "Tag (1-10)"
)

save(df_example_cli, file = "../data/df_example_cli.RData")
```

### Analysis

```{r}
library(franzpak)

df_example_cli <- df_example_cli |> rename(
  x = illtask, y = negativ, w = support
)

cortable_integriert2 <- cortable_multilevel(df_example_cli,
                                            varnames = c("w", "y", "x"),
                                            grp = "id")

cortable_integriert2 |> kableExtra::kable()
```

```{r}
ri.rs_modell <- lmer(y ~ x_dm + x_gm + y_dm_l1 + (1 + x_dm | id), data = df_example_cli)
model_parameters(ri.rs_modell) |> print_html(pretty_names = "labels")
```

```{r}
cli_model <- lmer(y ~ x_dm + x_gm + y_dm_l1 + w + w*x_dm + (1 + x_dm | id), data = df_example_cli)
model_parameters(cli_model) |> print_html(pretty_names = "labels")
```

```{r}
plot_model(cli_model, type = "int", terms = c("x_dm", "w"))
ggsave("Plots/examplepaper_CLI.png", width = 7, height = 5)
```

### Export to Excel table

```{r}
# cortable_integriert2 |> writexl::write_xlsx()
# h1_modelsummary <- model_parameters(ri.rs_modell)
# h2_modelsummary <- model_parameters(cli_model)
```
